---
title: "E1 Data analysis"
author: "Matt Crump"
date: "10/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

Data collected 2/10/22

## Load libraries

```{r}
library(pacman)
library(dplyr)
library(tidyverse)
library(jsonlite)
```

## Import Data

```{r}
# Read the text file from JATOS ...
read_file('data/E1/jatos_results_20220330174346.txt') %>%
  # ... split it into lines ...
  str_split('\n') %>% first() %>%
  # ... filter empty rows ...
  discard(function(x) x == '') %>%
  # ... parse JSON into a data.frame
  map_dfr(fromJSON, flatten=T) -> all_data
```

## Pre-processing

No subject exclusion

```{r}

```


# Accuracy analysis

## Get means across Conditions

```{r}
# select data from test phase
filtered_data <- all_data %>%
  filter(experiment_phase == "test")

# get mean accuracy in each condition for each subject
subject_means <- filtered_data %>%
  group_by(ID,
           encoding_stimulus_time,
           encoding_instruction,
           test_condition)%>%
  summarize(mean_correct = mean(correct))

# get mean accuracy in each condition across subjects
condition_means <- subject_means %>%
  group_by(encoding_stimulus_time,
           encoding_instruction,
           test_condition)%>%
  summarize(grp_mean_correct = mean(mean_correct),
            sem = sd(mean_correct)/sqrt(length(mean_correct)))

# print table
knitr::kable(condition_means)

# print graph
library(ggplot2)

e1_figure1 <- ggplot(condition_means, aes(x=encoding_instruction,y=grp_mean_correct,group=test_condition, color=test_condition))+
  geom_line()+
  geom_errorbar(aes(ymin = grp_mean_correct-sem,
                    ymax = grp_mean_correct+sem),
                width=.2)+
  geom_point()+
  facet_wrap(~encoding_stimulus_time)+
  coord_cartesian(ylim=c(.4,1))+
  geom_hline(yintercept=.5)+
  scale_y_continuous(breaks = seq(0.4,1,.1))

e1_figure1

```

## Main effect of test condition

```{r}

test_means <- subject_means %>%
  group_by(test_condition)%>%
  summarize(grp_mean_correct = mean(mean_correct),
            sem = sd(mean_correct)/sqrt(length(mean_correct)))

ggplot(test_means, aes(x=test_condition,y=grp_mean_correct))+
  geom_bar(stat="identity")+
  geom_errorbar(aes(ymin = grp_mean_correct-sem,
                    ymax = grp_mean_correct+sem),
                width=.2)+
  geom_point()+
  coord_cartesian(ylim=c(.4,1))+
  geom_hline(yintercept=.5)+
  scale_y_continuous(breaks = seq(0.4,1,.1))

```

## Main effect of stimulus duration

```{r}

duration_means <- subject_means %>%
  group_by(encoding_stimulus_time)%>%
  summarize(grp_mean_correct = mean(mean_correct),
            sem = sd(mean_correct)/sqrt(length(mean_correct)))

ggplot(duration_means, aes(x=encoding_stimulus_time,y=grp_mean_correct))+
  geom_bar(stat="identity")+
  geom_errorbar(aes(ymin = grp_mean_correct-sem,
                    ymax = grp_mean_correct+sem),
                width=.2)+
  geom_point()+
  coord_cartesian(ylim=c(.4,1))+
  geom_hline(yintercept=.5)+
  scale_y_continuous(breaks = seq(0.4,1,.1))

```

## Main effect of directed forgetting

```{r}

instruction_means <- subject_means %>%
  group_by(encoding_instruction)%>%
  summarize(grp_mean_correct = mean(mean_correct),
            sem = sd(mean_correct)/sqrt(length(mean_correct)))

ggplot(instruction_means, aes(x=encoding_instruction,y=grp_mean_correct))+
  geom_bar(stat="identity")+
  geom_errorbar(aes(ymin = grp_mean_correct-sem,
                    ymax = grp_mean_correct+sem),
                width=.2)+
  geom_point()+
  coord_cartesian(ylim=c(.4,1))+
  geom_hline(yintercept=.5)+
  scale_y_continuous(breaks = seq(0.4,1,.1))

```

## 2-way encoding_instruction x encoding_stimulus_time

```{r}
instruction_time <- subject_means %>%
  group_by(encoding_stimulus_time,encoding_instruction)%>%
  summarize(grp_mean_correct = mean(mean_correct),
            sem = sd(mean_correct)/sqrt(length(mean_correct)))

ggplot(instruction_time, aes(x=encoding_stimulus_time,
                            y=grp_mean_correct,
                            group=encoding_instruction,
                            fill=encoding_instruction))+
  geom_bar(stat="identity", position="dodge")+
  geom_errorbar(aes(ymin = grp_mean_correct-sem,
                    ymax = grp_mean_correct+sem),
                width=.2,
                position=position_dodge(width=0.9))+
  coord_cartesian(ylim=c(.4,1))+
  geom_hline(yintercept=.5)+
  scale_y_continuous(breaks = seq(0.4,1,.1))

```


## ANOVA

```{r}

subject_means$ID <- as.factor(subject_means$ID)

subject_means$encoding_stimulus_time <- as.factor(subject_means$encoding_stimulus_time)

subject_means$encoding_instruction <- as.factor(subject_means$encoding_instruction)

subject_means$test_condition <- as.factor(subject_means$test_condition)

aov.out <- aov(mean_correct ~ encoding_stimulus_time*encoding_instruction*test_condition + Error(ID/(encoding_stimulus_time*encoding_instruction*test_condition)), subject_means)

summary(aov.out)

```

## Write-up

```{r}
library(papaja)

E1_ANOVA <- papaja::apa_print(aov.out)
E1_ANOVA$full_result$encoding_stimulus_time

E1_ANOVA$full_result$test_condition

E1_ANOVA$full_result$encoding_instruction

E1_ANOVA$full_result$encoding_stimulus_time_encoding_instruction

## simple effects
IT500 <- subject_means %>%
  filter(encoding_stimulus_time == "500") %>%
  group_by(ID,encoding_instruction) %>%
  summarize(grp_mean_correct = mean(mean_correct))


IT500_report <- apa_print(t.test(IT500[IT500$encoding_instruction == "R",]$grp_mean_correct - IT500[IT500$encoding_instruction == "F",]$grp_mean_correct ))
IT500_report$full_result

IT1000 <- subject_means %>%
  filter(encoding_stimulus_time == "1000") %>%
  group_by(ID,encoding_instruction) %>%
  summarize(grp_mean_correct = mean(mean_correct))

IT1000_report <- apa_print(t.test(IT1000[IT1000$encoding_instruction == "R",]$grp_mean_correct - IT1000[IT1000$encoding_instruction == "F",]$grp_mean_correct))
IT1000_report$full_result

IT2000 <- subject_means %>%
  filter(encoding_stimulus_time == "2000") %>%
  group_by(ID,encoding_instruction) %>%
  summarize(grp_mean_correct = mean(mean_correct))

IT2000_report <- apa_print(t.test(IT2000[IT2000$encoding_instruction == "R",]$grp_mean_correct - IT2000[IT2000$encoding_instruction == "F",]$grp_mean_correct))
IT2000_report$full_result

```


Proportion correct for each subject in each condition was submitted to a 3 (Encoding Time: 500ms, 1000ms, 2000ms) x 2 (Encoding Instruction: Forget vs. Remember) x 2 (Lure type: Novel vs. Exemplar) fully repeated measures ANOVA. For completeness, each main effect and higher-order interaction is described in turn.

The main effect of encoding time was, `r E1_ANOVA$full_result$encoding_stimulus_time`. Proportion correct was `r round(duration_means[duration_means$encoding_stimulus_time == "500",]$grp_mean_correct, digits=3)` (`r round(duration_means[duration_means$encoding_stimulus_time == "500",]$sem,digits=3)`), `r round(duration_means[duration_means$encoding_stimulus_time == "1000",]$grp_mean_correct, digits=3)` (`r round(duration_means[duration_means$encoding_stimulus_time == "1000",]$sem,digits=3)`), and `r round(duration_means[duration_means$encoding_stimulus_time == "2000",]$grp_mean_correct, digits=3)` (`r round(duration_means[duration_means$encoding_stimulus_time == "2000",]$sem,digits=3)`) for the 500ms, 1000ms, and 2000ms intervals, respectively.

The main effect of test lure was significant, `r E1_ANOVA$full_result$test_condition`. Proportion correct was higher for novel lures (`r round(test_means[test_means$test_condition == "novel",]$grp_mean_correct, digits=3)`, SEM = `r round(test_means[test_means$test_condition == "novel",]$sem, digits=3)`), compared to exemplar lures (`r round(test_means[test_means$test_condition == "exemplar",]$grp_mean_correct, digits=3)`, SEM = `r round(test_means[test_means$test_condition == "exemplar",]$sem, digits=3)`).

The main effect of encoding instruction was, `r E1_ANOVA$full_result$encoding_instruction`. Proportion correct was similar for forget cued (`r round(instruction_means[instruction_means$encoding_instruction == "F",]$grp_mean_correct, digits=3)`, SEM = `r round(instruction_means[instruction_means$encoding_instruction == "F",]$sem, digits=3)`) and remember cued (`r round(instruction_means[instruction_means$encoding_instruction == "R",]$grp_mean_correct, digits=3)`, SEM = `r round(instruction_means[instruction_means$encoding_instruction == "R",]$sem, digits=3)`) items

The interaction between encoding instruction and encoding time was significant, `r E1_ANOVA$full_result$encoding_stimulus_time_encoding_instruction`. To further interpret this interaction, paired sample t-tests were used to assess the directed forgetting effect at each encoding time duration. The directed forgetting effect is taken as the difference between proportion correct for remember minus forget items. At 500 ms, the directed forgetting effect was not detected, `r IT500_report$full_result`. At 1000ms, the directed forgetting effect was reversed, `r IT1000_report$full_result`. And, at 2000 ms, the directed forgetting effect was again not detected, `r IT2000_report$full_result`. The remaining interactions were not significant. 

# RT analysis

## Get means across Conditions

```{r}
# select data from test phase
filtered_data <- all_data %>%
  filter(experiment_phase == "test",
         correct == TRUE)

# get mean accuracy in each condition for each subject
subject_means <- filtered_data %>%
  group_by(ID,
           encoding_stimulus_time,
           encoding_instruction,
           test_condition)%>%
  summarize(mean_RT = mean(rt))

# get mean accuracy in each condition across subjects
condition_means <- subject_means %>%
  group_by(encoding_stimulus_time,
           encoding_instruction,
           test_condition)%>%
  summarize(grp_mean_RT = mean(mean_RT),
            sem = sd(mean_RT)/sqrt(length(mean_RT)))

# print table
knitr::kable(condition_means)

# print graph
library(ggplot2)

e1_figure1_rt <- ggplot(condition_means, aes(x=encoding_instruction,y=grp_mean_RT,group=test_condition, color=test_condition))+
  geom_line()+
  geom_errorbar(aes(ymin = grp_mean_RT-sem,
                    ymax = grp_mean_RT+sem),
                width=.2)+
  geom_point()+
  facet_wrap(~encoding_stimulus_time)+
  coord_cartesian(ylim=c(1000,2000))+
  scale_y_continuous(breaks = seq(0,3000,250))

e1_figure1_rt

```

## Main effect of test condition

```{r}

test_means_rt <- subject_means %>%
  group_by(test_condition)%>%
  summarize(grp_mean_RT = mean(mean_RT),
            sem = sd(mean_RT)/sqrt(length(mean_RT)))

ggplot(test_means, aes(x=test_condition,y=grp_mean_RT))+
  geom_bar(stat="identity")+
  geom_errorbar(aes(ymin = grp_mean_RT-sem,
                    ymax = grp_mean_RT+sem),
                width=.2)+
  geom_point()+
  coord_cartesian(ylim=c(1000,2000))+
  scale_y_continuous(breaks = seq(0,3000,250))


```

## Main effect of stimulus duration

```{r}

duration_means_rt <- subject_means %>%
  group_by(encoding_stimulus_time)%>%
  summarize(grp_mean_RT = mean(mean_RT),
            sem = sd(mean_RT)/sqrt(length(mean_RT)))

ggplot(duration_means, aes(x=encoding_stimulus_time,y=grp_mean_RT))+
  geom_bar(stat="identity")+
  geom_errorbar(aes(ymin = grp_mean_RT-sem,
                    ymax = grp_mean_RT+sem),
                width=.2)+
  geom_point()+
  coord_cartesian(ylim=c(1000,2000))+
  scale_y_continuous(breaks = seq(0,3000,250))

```

## Main effect of directed forgetting

```{r}

instruction_means <- subject_means %>%
  group_by(encoding_instruction)%>%
  summarize(grp_mean_RT = mean(mean_RT),
            sem = sd(mean_RT)/sqrt(length(mean_RT)))

ggplot(instruction_means, aes(x=encoding_instruction,y=grp_mean_RT))+
  geom_bar(stat="identity")+
  geom_errorbar(aes(ymin = grp_mean_RT-sem,
                    ymax = grp_mean_RT+sem),
                width=.2)+
  geom_point()+
  coord_cartesian(ylim=c(1000,2000))+
  scale_y_continuous(breaks = seq(0,3000,250))

```

## 2-way encoding_instruction x encoding_stimulus_time

```{r}
instruction_time <- subject_means %>%
  group_by(encoding_stimulus_time,encoding_instruction)%>%
  summarize(grp_mean_RT = mean(mean_RT),
            sem = sd(mean_RT)/sqrt(length(mean_RT)))

ggplot(instruction_time, aes(x=encoding_stimulus_time,
                            y=grp_mean_RT,
                            group=encoding_instruction,
                            fill=encoding_instruction))+
  geom_bar(stat="identity", position="dodge")+
  geom_errorbar(aes(ymin = grp_mean_RT-sem,
                    ymax = grp_mean_RT+sem),
                width=.2,
                position=position_dodge(width=0.9))+
  coord_cartesian(ylim=c(1000,2000))+
  scale_y_continuous(breaks = seq(0,3000,250))

```


## ANOVA

```{r}

subject_means$ID <- as.factor(subject_means$ID)

subject_means$encoding_stimulus_time <- as.factor(subject_means$encoding_stimulus_time)

subject_means$encoding_instruction <- as.factor(subject_means$encoding_instruction)

subject_means$test_condition <- as.factor(subject_means$test_condition)

aov.out_rt <- aov(mean_RT ~ encoding_stimulus_time*encoding_instruction*test_condition + Error(ID/(encoding_stimulus_time*encoding_instruction*test_condition)), subject_means)

summary(aov.out_rt)

```

## Write-up

```{r}
library(papaja)

E1_ANOVA_RT <- papaja::apa_print(aov.out_rt)
E1_ANOVA_RT$full_result$encoding_stimulus_time

E1_ANOVA_RT$full_result$test_condition

E1_ANOVA_RT$full_result$encoding_instruction

E1_ANOVA_RT$full_result$encoding_stimulus_time_encoding_instruction

## simple effects
IT500 <- subject_means %>%
  filter(encoding_stimulus_time == "500") %>%
  group_by(ID,encoding_instruction) %>%
  summarize(grp_mean_RT = mean(mean_RT))


IT500_report <- apa_print(t.test(IT500[IT500$encoding_instruction == "R",]$grp_mean_RT - IT500[IT500$encoding_instruction == "F",]$grp_mean_RT ))
IT500_report$full_result

IT1000 <- subject_means %>%
  filter(encoding_stimulus_time == "1000") %>%
  group_by(ID,encoding_instruction) %>%
  summarize(grp_mean_RT = mean(mean_RT))

IT1000_report <- apa_print(t.test(IT1000[IT1000$encoding_instruction == "R",]$grp_mean_RT - IT1000[IT1000$encoding_instruction == "F",]$grp_mean_RT))
IT1000_report$full_result

IT2000 <- subject_means %>%
  filter(encoding_stimulus_time == "2000") %>%
  group_by(ID,encoding_instruction) %>%
  summarize(grp_mean_RT = mean(mean_RT))

IT2000_report <- apa_print(t.test(IT2000[IT2000$encoding_instruction == "R",]$grp_mean_RT - IT2000[IT2000$encoding_instruction == "F",]$grp_mean_RT))
IT2000_report$full_result

```


Mean reaction times for each subject in each condition was submitted to a 3 (Encoding Time: 500ms, 1000ms, 2000ms) x 2 (Encoding Instruction: Forget vs. Remember) x 2 (Lure type: Novel vs. Exemplar) fully repeated measures ANOVA. For completeness, each main effect and higher-order interaction is described in turn.

The main effect of encoding time was, `r E1_ANOVA$full_result$encoding_stimulus_time`. Proportion correct was `r round(duration_means[duration_means$encoding_stimulus_time == "500",]$grp_mean_correct, digits=3)` (`r round(duration_means[duration_means$encoding_stimulus_time == "500",]$sem,digits=3)`), `r round(duration_means[duration_means$encoding_stimulus_time == "1000",]$grp_mean_correct, digits=3)` (`r round(duration_means[duration_means$encoding_stimulus_time == "1000",]$sem,digits=3)`), and `r round(duration_means[duration_means$encoding_stimulus_time == "2000",]$grp_mean_correct, digits=3)` (`r round(duration_means[duration_means$encoding_stimulus_time == "2000",]$sem,digits=3)`) for the 500ms, 1000ms, and 2000ms intervals, respectively.

The main effect of test lure was significant, `r E1_ANOVA$full_result$test_condition`. Proportion correct was higher for novel lures (`r round(test_means[test_means$test_condition == "novel",]$grp_mean_correct, digits=3)`, SEM = `r round(test_means[test_means$test_condition == "novel",]$sem, digits=3)`), compared to exemplar lures (`r round(test_means[test_means$test_condition == "exemplar",]$grp_mean_correct, digits=3)`, SEM = `r round(test_means[test_means$test_condition == "exemplar",]$sem, digits=3)`).

The main effect of encoding instruction was, `r E1_ANOVA$full_result$encoding_instruction`. Proportion correct was similar for forget cued (`r round(instruction_means[instruction_means$encoding_instruction == "F",]$grp_mean_correct, digits=3)`, SEM = `r round(instruction_means[instruction_means$encoding_instruction == "F",]$sem, digits=3)`) and remember cued (`r round(instruction_means[instruction_means$encoding_instruction == "R",]$grp_mean_correct, digits=3)`, SEM = `r round(instruction_means[instruction_means$encoding_instruction == "R",]$sem, digits=3)`) items

The interaction between encoding instruction and encoding time was significant, `r E1_ANOVA$full_result$encoding_stimulus_time_encoding_instruction`. To further interpret this interaction, paired sample t-tests were used to assess the directed forgetting effect at each encoding time duration. The directed forgetting effect is taken as the difference between proportion correct for remember minus forget items. At 500 ms, the directed forgetting effect was not detected, `r IT500_report$full_result`. At 1000ms, the directed forgetting effect was reversed, `r IT1000_report$full_result`. And, at 2000 ms, the directed forgetting effect was again not detected, `r IT2000_report$full_result`. The remaining interactions were not significant. 

## Attention check

Did individual participants successfully press the F or R keys during encoding?

```{r}
# select data from test phase
filtered_data <- all_data %>%
  filter(experiment_phase == "study",
         is.na(correct) == FALSE)

# get mean accuracy in each condition for each subject
subject_means <- filtered_data %>%
  group_by(ID)%>%
  summarize(mean_correct = mean(correct))

knitr::kable(subject_means)

```


## Design

```{r}
sub_1 <- all_data %>%
  filter(ID =="j1s5gca39ee2lkr9",
         item_type == "study",
         experiment_phase == "study") %>%
  group_by(category_type,category,encoding_instruction,stimulus_time) %>%
  count()

```

## Demographics

```{r}
library(tidyr)

demographics <- all_data %>%
  filter(trial_type == "survey-html-form") %>%
  select(ID,response) %>%
  unnest_wider(response) %>%
  mutate(age = as.numeric(age))

age_demographics <- demographics %>%
  summarize(mean_age = mean(age),
            sd_age = sd(age),
            min_age = min(age),
            max_age = max(age))

factor_demographics <- apply(demographics[-1], 2, table)

```

A total of `r dim(demographics)[1]` participants were recruited from Amazon's Mechanical Turk. Mean age was `r round(age_demographics$mean_age, digits=1)` (range = `r age_demographics$min_age` to `r age_demographics$max_age` ). There were `r as.numeric(factor_demographics$sex["female"])` females, and `r as.numeric(factor_demographics$sex["male"])` males. There were `r as.numeric(factor_demographics$hand["Right"])` right-handed participants, and `r as.numeric(factor_demographics$hand["Both"])+as.numeric(factor_demographics$hand["Left"])` left or both handed participants. `r as.numeric(factor_demographics$vision["Normal"])` participants reported normal vision, and `r as.numeric(factor_demographics$vision["Corrected"])` participants reported corrected-to-normal vision. `r as.numeric(factor_demographics$english["First"])` participants reported english as a first language, and `r as.numeric(factor_demographics$english["Second"])` participants reported english as a second language.

## save environment

```{r}
save.image("data/E1/E1_data.RData")
```


