---
title: "Pilot analysis"
author: "Matt Crump"
date: "10/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is a pilot analysis of 5 subjects worth of data collected from a pilot version of this experiment. Data collected 12/2021.

## Load libraries

```{r}
library(pacman)
library(dplyr)
library(tidyverse)
library(jsonlite)
```



## Import Data

```{r}
# Read the text file from JATOS ...
read_file('jatos_results_20211206204357.txt') %>%
  # ... split it into lines ...
  str_split('\n') %>% first() %>%
  # ... filter empty rows ...
  discard(function(x) x == '') %>%
  # ... parse JSON into a data.frame
  map_dfr(fromJSON, flatten=T) -> data
```

## Pre-processing

Note: In the pilot experiment script, the stimulus duration time was encoded to the data file during the encoding phase, but not during the recognition memory test phase. The following R code chunk, analyzes the encoding phase for each participant, and adds the stimulus duration time factor to the recognition memory test phase data.

```{r}

# add stimulus encoding factor

all_data <- data.frame()
for( sub in unique(data$ID)){
  
  temp_data <- data %>%
    filter(ID==sub)
  
  temp_study <- temp_data %>%
    filter(experiment_phase == "study")
  
  temp_test <- temp_data %>%
    filter(experiment_phase == "test")
  
  encoding_time <- c()
  for(i in 1: dim(temp_test)[1]){
    im_name <- temp_test[i,]$study_item_image_name
    get_trial <- temp_study %>%
      filter(image_name == im_name)
    encoding_time[i]<-get_trial$stimulus_time
  }
  
  temp_test <- cbind(temp_test,encoding_time)
  all_data <- rbind(all_data,temp_test)
}
```

## Determine encoding performance

```{r}
encoding_data <- data %>%
  filter(experiment_phase == "study",
         is.na(correct) == FALSE )

# get mean accuracy in each condition for each subject
get_means <- encoding_data %>%
  group_by(ID)%>%
  summarize(mean_correct = mean(correct))

knitr::kable(get_means)

```

## Get means across Conditions

```{r}
# select data from test phase
filtered_data <- all_data %>%
  filter(experiment_phase == "test")

# get mean accuracy in each condition for each subject
get_means <- filtered_data %>%
  group_by(ID,encoding_time,encoding_instruction,test_condition)%>%
  summarize(mean_correct = mean(correct))

# get mean accuracy in each condition across subjects
get_group_means <- get_means %>%
  group_by(encoding_time,encoding_instruction,test_condition)%>%
  summarize(grp_mean_correct = mean(mean_correct),
            sem = sd(mean_correct)/sqrt(length(mean_correct)))

# print table
knitr::kable(get_group_means)

# print graph
library(ggplot2)

ggplot(get_group_means, aes(x=encoding_instruction,y=grp_mean_correct,group=test_condition, color=test_condition))+
  geom_line()+
  geom_errorbar(aes(ymin = grp_mean_correct-sem,
                    ymax = grp_mean_correct+sem),
                width=.2)+
  geom_point()+
  facet_wrap(~encoding_time)+
  coord_cartesian(ylim=c(.4,1))+
  geom_hline(yintercept=.5)+
  scale_y_continuous(breaks = seq(0.4,1,.1))

```

## Main effect of test condition

```{r}

get_group_means <- get_means %>%
  group_by(test_condition)%>%
  summarize(grp_mean_correct = mean(mean_correct),
            sem = sd(mean_correct)/sqrt(length(mean_correct)))

ggplot(get_group_means, aes(x=test_condition,y=grp_mean_correct))+
  geom_bar(stat="identity")+
  geom_errorbar(aes(ymin = grp_mean_correct-sem,
                    ymax = grp_mean_correct+sem),
                width=.2)+
  geom_point()+
  coord_cartesian(ylim=c(.4,1))+
  geom_hline(yintercept=.5)+
  scale_y_continuous(breaks = seq(0.4,1,.1))

```

## Main effect of stimulus duration

```{r}

get_group_means <- get_means %>%
  group_by(encoding_time)%>%
  summarize(grp_mean_correct = mean(mean_correct),
            sem = sd(mean_correct)/sqrt(length(mean_correct)))

ggplot(get_group_means, aes(x=encoding_time,y=grp_mean_correct))+
  geom_bar(stat="identity")+
  geom_errorbar(aes(ymin = grp_mean_correct-sem,
                    ymax = grp_mean_correct+sem),
                width=.2)+
  geom_point()+
  coord_cartesian(ylim=c(.4,1))+
  geom_hline(yintercept=.5)+
  scale_y_continuous(breaks = seq(0.4,1,.1))

```

## Main effect of directed forgetting

```{r}

get_group_means <- get_means %>%
  group_by(encoding_instruction)%>%
  summarize(grp_mean_correct = mean(mean_correct),
            sem = sd(mean_correct)/sqrt(length(mean_correct)))

ggplot(get_group_means, aes(x=encoding_instruction,y=grp_mean_correct))+
  geom_bar(stat="identity")+
  geom_errorbar(aes(ymin = grp_mean_correct-sem,
                    ymax = grp_mean_correct+sem),
                width=.2)+
  geom_point()+
  coord_cartesian(ylim=c(.4,1))+
  geom_hline(yintercept=.5)+
  scale_y_continuous(breaks = seq(0.4,1,.1))

```
