---
title: "Directed Forgetting Pictures 1: Stimulus presentation time"
author: "Patrick Ihejirika and Matt Crump"
date: "8/16/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

This file functions as a readme and sketchpad for experiment design during project development

Ahmad, F. N., Tan, P., & Hockley, W. E. (2019). Directed forgetting for categorised pictures: Recognition memory for perceptual details versus gist. Memory, 27(7), 894–903. https://doi.org/10/gmgg3g

This experiment script attempts to closely replicate the design of Experiment 1.

Changes will be listed here:

downloaded natural scenes from Brady
edited file names to lowercase jpg

## testing out the idea of random factors

```{r}

stimulus_times <- runif(n = 100*120, min = .1, max = 5)
cue_times <- runif(n = 100*120, min = .5, max = 5)

trials <- data.frame(stimulus_times,cue_times)

plot(stimulus_times,cue_times)

```


```{r}
#load sample data
library(dplyr)
a<- jsonlite::fromJSON("test_json.txt")

#check encoding trials have equal F and R trials per category
a %>%
  filter(experiment_phase == "study")%>%
  group_by(category_type,category,encoding_instruction,stimulus_time) %>%
  count()

filtered <-a %>% filter(experiment_phase %in% c("study","test")) %>%
  select(stimulus,study_item_image_name,test_item_image_name)

```



```{r}

remember <- rbinom(60,1,prob=.9)
forget <- rbinom(60,1,prob=.9-.07)

sum(rbinom(60,1,prob=.9))/60
sum(rbinom(60,1,prob=.83))/60

get_df_effect <- function(r,f){
(sum(rbinom(60,1,prob=r))/60) - (sum(rbinom(60,1,prob=f))/60)
}

get_df_effect(.9,.83)

hist(replicate(50,get_df_effect(.9,.89)))

```

